Hydras \& Co.

### 2022-03-17 doc/part-hydras.tex
corrections chapter 5 (up to p. 106)

\begin{itemize}
\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and a fortiori 
$[0,\omega)$ (section ~\vref{omega-case}).

@@

Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
come from 
 an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
 by L.~Kirby and J.~Paris.% on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
 Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
rounds of hydra battles. 
Nevertheless, they have the same global structure as the simple proofs described in
sections~\vref{omega-case} and \vref{omega2-case}. 
We invite the reader to compare the three proofs step by step, lemma by lemma.

@@

\subsubsection{Canonical sequences in Coq}
\index{hydras}{Library Epsilon0!Functions!canon}
\index{hydras}{Library Epsilon0!Functions!canonS}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}

@@

@@

We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}

For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.

@@

We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i+1}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$). This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.

\label{lemma:canonS_LT}

\input{movies/snippets/Canon/canonSLT}

@@

@@

\noindent 
From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we can convert any path into an inequality on ordinals (by induction on paths).

>>>>

\begin{itemize}
\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).

@@

Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
come from 
 an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
 by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
 Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
rounds of hydra battles. 
Nevertheless, they have the same global structure as the simple proofs described in
sections~\vref{omega-case} and \vref{omega2-case}. 
We invite the reader to compare the three proofs step by step, lemma by lemma.

@@

\subsubsection{Canonical sequences in \coq}
\index{hydras}{Library Epsilon0!Functions!canon}
\index{hydras}{Library Epsilon0!Functions!canonS}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}

@@

\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}.

@@

We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.

@@

We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$). This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.

\label{lemma:canon_LT}

\inputsnippets{Canon/canonLT}

\paragraph*{\gaiasign} This lemma is also available in
Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:

\inputsnippets{GCanon/gcanonLt}

@@

\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
\inputsnippets{GCanon/gcanonLimitOf}

@@

\noindent 
From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
